Collaborative Networks in Machine Learning Hardware\\ \thanks{University of Texas at Austin}

### 2024-03-11 paper/paper.tex
prelim fix of tables, added conclusion

\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
% mine own
\usepackage{caption}
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
\begin{document}

\title{Conference Paper Title\\
\thanks{Identify applicable funding agency here. If none, delete this.}
}

@@

In finding additional outliers, we were also able to distinguish the age of our papers. 
The dataset consisted of papers from the years of 1971 to 2023. Since machine 
learning was identified as a rising field from the discussion regarding Figure 2, 
there were many papers from 2023. And, since many papers only provided a month as 
the date of publication, the multiple papers from November of 2023 prevented the 
identification of the newest paper. However, the oldest paper, the only paper from 
1971, was titled “Unit Spike Activity in Coelenteran Neural Network”, a paper about 
temperature spikes in primitive neural nerve networks [10]. \par

@@

\begin{figure}[htbp]
    % \centering
    \caption{Number of Papers per Year by Category}
    \captionsetup{labelformat=empty}
    \parbox[b]{.5\linewidth}{
    \includegraphics[scale=0.23]{papers_by_year.png}
    \caption{A cat}\label{cat}}
    \parbox[b]{.45\linewidth}{
    \includegraphics[scale=0.23]{times_cited.png}
    \caption{An elephant}\label{elephant}}
%   \caption{Number of Papers per Year by Category}\label{animals}
    \captionsetup{labelformat=simple}
\end{figure}

In Table 1, we examine the structure of each community by computing the mixed 
edge probability, i.e. how many out of the total edges in the graph contain 
an edge from one community to other communities. The mixed edge theoretical 
probability of a certain category, say A, is the probability that any node 
is in A out of all nodes, multiplied by the probability that any node is 
in any other category. We use this simple equation: \par

% TODO: fix new line problem!
\begin{equation}
    Theoretical\; Mixed\; Edge\; Probability(category\; A)
    \\
    = (P(node\; in\;A)) (1 - P(node\;in\;A))\label{eq}
\end{equation}

>>>>

\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
% mine own
% \usepackage{caption}
\usepackage{tabularx}
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
\begin{document}

\title{Collaborative Networks in Machine Learning Hardware\\
\thanks{University of Texas at Austin}
}

@@

In finding additional outliers, we were also able to distinguish the age of our papers. 
The dataset consisted of papers from the years of 1971 to 2023. The oldest paper, 
the only paper from 1971, was titled “Unit Spike Activity in Coelenteran Neural Network”, 
a paper about temperature spikes in primitive neural nerve networks [10].  \par

@@

\begin{figure}[htbp]
    \centering
    \caption{Number of Papers per Year by Category}
    % \captionsetup{labelformat=empty}
    \parbox[b]{.5\linewidth}{
    \includegraphics[scale=0.23]{papers_by_year.png}
    \caption{A cat}\label{cat}}
    \parbox[b]{.45\linewidth}{
    \includegraphics[scale=0.23]{times_cited.png}
    \caption{An elephant}\label{elephant}}
    \hfill
    \parbox[b]{.5\linewidth}{
    \includegraphics[scale=0.23]{papers_by_year.png}
    \caption{A cat}\label{cat}}
    \parbox[b]{.45\linewidth}{
    \includegraphics[scale=0.23]{times_cited.png}
    \caption{An elephant}\label{elephant}}
    % \captionsetup{labelformat=simple}
\end{figure}

Next, we examine the structure of each community by computing the mixed edge 
probability, i.e. how many out of the total edges in the graph contain an edge 
from one community to other communities. The mixed edge theoretical probability 
of a certain category, say A, is the probability that any node is in A out of 
all nodes, multiplied by the probability that any node is in any other category. 
We use this simple equation: \par

% TODO: fix new line problem!
\begin{equation}
    Theoretical\; Mixed\; Edge\; Probability(category\; A)
\end{equation}
\begin{equation}
    = (P(node\; in\;A)) (1 - P(node\;in\;A))\label{eq1}
\end{equation}

This number indicates the portion out of all edges that should contain an edge stretching 
from a node in A to a node outside of A if the graph picked edges at random. \par

Our actual mixed edge probability is calculated by: \par
\begin{equation}
    \dfrac{Number\; of\; in\; or\; out\; edges\; of\; a\; community}{Number\; of\; total\; edges}
    \label{eq2}
\end{equation}

As seen in Table 2 below, our actual mixed edge probability is far 
below the theoretical probability in a random graph. \par

% Todo: fix new line problem!
\begin{table}[htbp]
    \caption{Mixed Edge Categorical Analysis}
    \begin{center}
        \begin{tabular}{|c|cccc|}
        \hline
        \textbf{Categories} & \textbf{\textit{Edge}} & \textbf{\textit{ML}} & 
        \textbf{\textit{Neuromorphic}} & \textbf{\textit{Spiking}} \\
        \textbf{} & \textbf{\textit{ML}} & \textbf{\textit{Accelerator}} & 
        \textbf{\textit{Computing}} & \textbf{\textit{NN}} \\
        \hline
        \textbf{\textit{Theoretical}} &\ &\ &\ &\ \\
        \textbf{\textit{Mixed Edge}} &
        0.19813& 0.19256& 0.19051& 0.16663 \\
        \textbf{\textit{Probability}} & \ &\ &\ &\ \\
        \ & \ &\ &\ &\ \\
        \textbf{\textit{Mixed Edge}} & 
        0.02101& 0.06806& 0.09880& 0.05195 \\
        \textbf{\textit{Probability}} & \ &\ &\ &\ \\
        \hline
        \end{tabular}
        \label{tab2}
    \end{center}
\end{table}

% TODO: can be removed later
\ \newline
\ \newline
\ \newline
\ \newline
\ \newline
\ \newline

There are far fewer edges in between different categories than edges within a category. 
As such, these topics within machine learning are distinct subject areas. \par

\subsection{GNNs}

For our baseline, we used a basic five-layer neural network that is essentially trained on word 
vectors. On Cora, this achieved a 79.43\% accuracy for node classification; 
on Citeseer, this achieved a 71.43\% accuracy. We then implemented a graphical 
convolution network model on the Cora dataset with five feed-forward layers, which 
achieves an accuracy of 71.13\% for link prediction and an accuracy of  86.58\% for 
node classification. For the CiteSeer dataset, we received an accuracy of 74.1
\% on node classification and 84.29\% on link prediction. Similarly, the PubMed dataset 
yielded an accuracy of 78.7\% for both node classification and link prediction. 
These results were further verified through applying a GNN model from DGL, which received 
similar results to the previously mentioned models. This makes sense since Cora is a dataset 
of machine learning papers while CiteSeer has software engineering papers, which is a larger 
domain. Thus, citations are easier to predict in Citeseer. Furthermore, since Citeseer 
has as many keywords as they do papers, and there are half as many keywords in Cora, 
it is more likely that testing on Cora is done on keywords that the model has seen 
before and is thus more accurate. \par

In this regard, these datasets helped to establish a baseline for the expected performance 
of a GNN on the dataset created from the Web of Science after PCA. Following similar methods as 
before, the GNN model for node classification produced an accuracy of 78.06\%, 
while Link prediction fared much better with an accuracy of 82.56\%. 
These results are summarized in Table 3 below. \par

\begin{table}[htbp]
    \caption{GNN Performance Across Datasets}
    \begin{center}
        \begin{tabular}{|c|cccccc|}
        \hline
        \textbf{Dataset} & \textbf{\textit{Cora}} & \textbf{\textit{CS$^{\mathrm{a}}$}} & \textbf{\textit{PM$^{\mathrm{b}}$}} &
        \textbf{\textit{WoS1$^{\mathrm{c}}$}} & \textbf{WoS2$^{\mathrm{d}}$} & \textbf{WoS3$^{\mathrm{e}}$} \\
        \hline
        \textbf{\textit{Node}} &\ &\ &\ &\ &\ &\ \\
        \textbf{\textit{Classif-}} &
        86.58\%& 74.1\%& 78.7\%& 72.3\%& 76.96\%& 78.06\% \\
        \textbf{\textit{ication}} &\ &\ &\ &\ &\ &\ \\

        \ &\ &\ &\ &\ &\ &\ \\

        \textbf{\textit{Link}} &\ &\ &\ &\ &\ &\ \\
        \textbf{\textit{Predic-}} &
        71.13\%& 84.29\%& 78.74\%& 80.8\%& 81.77\%& 82.56\% \\
        \textbf{\textit{tion}} &\ &\ &\ &\ &\ &\ \\

        \hline
        \end{tabular}
        \label{tab1}
        {$^{\mathrm{a}}$CiteSeer.}
        {$^{\mathrm{b}}$PubMed.}
        {$^{\mathrm{c}}$Small WoS.}
        {$^{\mathrm{d}}$Large WoS.}
        {$^{\mathrm{e}}$Large WoS w/ PCA.}
    \end{center}
\end{table}

From the results in the table, we see that our dataset performs weaker than 
PubMed and Cora in terms of node classification and weaker than CiteSeer in 
terms of link prediction. Despite this, our model has the strongest performance 
averaged between node classification and link prediction, even before PCA. It is 
also interesting to note that the larger networks in terms of nodes have a smaller 
difference between the performance of node classification and link prediction, 
WoS by just a few percentage points and PubMed by less than a tenth of a percent. 
Similarly, CiteSeer has the lowest average degree and also performs the 
best for link prediction. \par

However, one of the strongest limitations in the results of the applied models 
is the lack of knowledge of the cause of performance. The performance of the 
models could be motivated by differences in the structure of the networks due 
to the nature of the associated fields, differences in the databases used to 
curate the networks, and differences in the characteristics of the networks 
themselves. For example, CiteSeer has multiple self-loops, while the other 
networks have few if any. Similarly, Cora and CiteSeer are unconnected, while 
PubMed is connected. Future research on this paper could extend this study to 
homogenize the datasets by removing self-loops, making all graphs connected, 
and applying PCA to all datasets to observe any changes in performance. \par

Similarly, a limitation in our approach exists in our dataset in the class 
imbalance of the ML Accelerator class having far fewer papers than the other 
classes. This limitation can influence the results of the dataset, such as 
creating a tendency to not classify a node as ML Accelerator since it is 
less likely for a paper to belong to this class. This issue can be addressed 
through solutions such as minority oversampling, creating artificial samples 
of the ML Accelerator class, or majority undersampling, omitting some samples 
of the remaining classes to simulate more balanced classes. However, to 
preserve any information about the underlying structure of the field of 
ML hardware, we assume that this imbalance of classes is a property of 
the field of ML hardware, avoiding any manipulation of the original 
set of samples. \par

For future plans, the most straightforward path is exploring the use of the 
dataset as an academic tool. Cora and CiteSeer are both extensively used as a 
tool for the exploration of machine learning, aiming for the strongest results 
of classification for the dataset. A similar approach can be used for the new 
WoS dataset, observing the results of these high-performing methods in Cora 
or CiteSeer on the WoS dataset. Similarly, this dataset can be used to explore 
new models for classifying or predicting citation networks. Since this dataset 
has different characteristics than all of the others, it may offer insight 
into using new or different models than before. \par

Additionally, the models trained on the WoS dataset can be used for node 
classification or link prediction in a practical sense. By classifying 
papers in the field of machine learning hardware, the model can help 
differentiate papers of the two categories in real life. This also 
applies to link prediction - predicting links on an unpublished paper or 
abstract can provide insight into potential inspiration from other academic 
works. Similar applications of link prediction can assist in finding not only 
papers for inspiration, but also other authors to collaborate with if, 
for example, both authors have similar published papers. \par

\section{Conclusion and Future Work}

Overall, we reached our goals for this project of developing a new dataset 
for papers relating to machine learning hardware for a better understanding 
of the relationships between publications and authors. Through creating a 
dataset that is well explained through node classification and link prediction, 
these results can be reproduced for alternative applications like publication 
classification or potential author collaboration. \par

In the development of this project, we have had the opportunity to witness 
the application of GNNs on citation networks, in addition to a new dataset. 
So, we have learned of various qualities of citation networks such as their 
degree distribution, clustering coefficient, and their underlying network 
structure, witnessing underlying paper data distribution. We also explore 
GNN model differences between node classification and link prediction. 
Creating a new dataset, we saw that GNNs are effective for node classification 
and link prediction on a new network of machine learning hardware papers. \par

Thus far, we created a dataset of machine learning hardware papers across 
different demographics with classifications to be used in node classification 
and link prediction, quantified in terms of characteristics like degree and 
clustering coefficient, comparing the datasets in terms of traits and GNN performance. 
While the creation of a competing academic dataset is a strong start to our goal, 
our created graph still has opportunities for improvement. \par
